Hierarchical wrinkling in a confined permeable biogel

### 2015-02-12 fig_plis/article_plis.tex
SI dans Dynamics of the wrinkling process

To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy. As shown in Supplementary Figure~\ref{fig:nonmonotonic} the gel consist in a space spanning fractal network with a cut off length scale $\xi$ that corresponds to the maximum pore size of the network. This model fits the structure at all times but its parameters vary non monotonously during GDL acidification. As shown in Figure~\ref{fig:acidification}d, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network [refXXX]. This stiffens the arms of the network and is responsible for the increase of elasticity. Beyond the isoelectric point $\xi$ decreases down to \SI{4}{\micro\metre}. As casein regain solubility, the strands of the network swell, the pores shrink and the gel weakens.

@@

Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favourable to wrinkling. Up to 10 min the dispersion is homogeneous and fills the entire slit. The black margins correspond to the glass walls. Around 23 min, the gel forms and simultaneously becomes thinner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the situation at 10min. This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides. We describe sample as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b). Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods). As expected from the system design, the evolution of the volume is non-monotonous. After reaching a minimum around 35\% of its initial value at 23 min, the volume increases. Shortly after,  at 36 min, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area. This confirms that the instability is due to swelling. Around 38 min, the amplitude of the wrinkles reach the thickness of the slit. The size of the primary pattern is now settled Supplementary Fig. \ref{fig:patterns, fig:growth}. However swelling continues, the gel flattens on both top and bottom walls (44 min) and finally buckles back (53 min) to form the secondary pattern (1h21). When the excess volume saturates, the dynamics stops. The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we have observe the formation of higher order patterns.

>>>>

To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy. As shown in Supplementary Figure~\ref{fig:nonmonotonic} the gel consist in a space spanning fractal network with a cut off length scale $\xi$ that corresponds to the maximum pore size of the network. This model fits the structure at all times but its parameters vary non monotonously during GDL acidification. As shown in Figure~\ref{fig:acidification}d, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network [refXXX]. This stiffens the arms of the network and is responsible for the increase of elasticity. Beyond the isoelectric point $\xi$ decreases down to \SI{4}{\micro\metre}. As caseins regain solubility, the strands of the network swell, the pores shrink and the gel weakens.

@@

Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favourable to wrinkling (see also Supplementary Videos 7 and 8). Initially the dispersion is homogeneous and fills the entire slit. The black margins correspond to the glass walls. Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation. This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides. We describe sample as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b). Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods). As expected from the system design, the evolution of the volume is non-monotonous. After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases. Shortly after,  at \SI{36}{\minute}, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area. This confirms that the instability is due to swelling. Around \SI{38}{\minute}, the amplitude of the wrinkles reach the thickness of the slit. The size of the primary pattern is now settled, Supplementary Figures~\ref{fig:patterns} and \ref{fig:growth}. However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (1h21). When the excess volume saturates, the dynamics stops. The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we have observe the formation of higher order patterns.
